---
title: "Draft Presentation"
author: "Ethan Schilling, Ivan Shokhrin, & Tyler Manuello"
date: "2025-12-01"
output:
  slidy_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Goal: Recreate a heatmapped Correlation matrix of 2018 NBA data and compare to the 2022 season
- Use a confusion/correlation to predict/understand All-NBA Team Placement status
- Group top 15 players of 2022 season and compare to 2018 players from the paper "A Hybrid Machine Learning Model for Predicting USA NBA All-Stars" 
- Use Monte Carlo and Bootstrapping to produce averaged estimates about NBA stats and their correlations.


## Introduction to Basic Stats
- Percent stats
  - 3pt, Ft, Fg
- Per game stats
  - Blocks, Steals, Assists, Rebounds, Points, Minutes
- Attempts
  - 3pt, Ft, Fg
- Total
  - Games Played

## What We Notice

```{r}
tibble1 <- readRDS("league_compare")
tibble1
```
- Pace is down
- Less turnovers
- Rebounding is up
- Smaller sample size
