---
title: "Correlation Of NBA Statistics"
author: "Ethan Schilling, Ivan Shokhrin, & Tyler Manuello"
date: "2025-12-01"
output: 
  ioslides_presentation:
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Goal: Understand Correlations between NBA performance stats
- Recreate Heatmapped Correlation Matrix from the article
- Quantify uncertainty using bootstrapping
- Simulate season-to-season variability with Monte Carlo
- Discuss important stats and features of the leauge that would lead to differing correlations across seasons.

##  Data and Methods

* **Dataset:** Top 15 NBA players (2018/2022 season stats)
* **Goal:** Understand relationships between NBA performance metrics

###  Methods Used
* Correlation analysis
* Bootstrap confidence intervals
* Monte Carlo simulation of correlation matrices


## Introduction to Basic Stats (talk about why bootstrapping is preferable for skewed variables, some are skewed)
- Percent stats
  - 3pt, Ft, Fg
- Per game stats
  - Blocks, Steals, Assists, Rebounds, Points, Minutes
- Attempts
  - 3pt, Ft, Fg
- Total
  - Games Played


## What We Notice from Season to Season
```{r}
tibble1 <- readRDS("league_compare")
tibble1
```
- Pace is down
- Less turnovers
- Rebounding is up
- Smaller sample size


## Replicated Correlation Matrix from 2018 season
```{r, fig.width=8, fig.height=4.5, fig.align='center'}
corr18 <- readRDS("corr18.rds")
corr18
```


## Statistics to Note from Matrix:
The red squares are the least correlated statistics where as the green squares are highly correlated on a scale from -1 to 1.

### Strong Relationships:
- PTS and FGA
- AST and PTS
- REB and BLK

### Weak relationships
- FT% and every other statistic


Maybe talk about why correlations would vary based on position on team


## Bootstrapping Correlations
Include bootstrap plot,
Table of CI's

Plot Ideas:

Simulated Correlation Histogram

Example: distribution of PTS–FGA correlation over 10,000 fake seasons.

Simulated Average Correlation Matrix vs Real Matrix

Two heatmaps side by side

Shows how robust the relationships are

Monte Carlo Variability Plot

For each stat pair, plot SD of simulated correlations

Identifies stable vs unstable relationships

Talking points (3 minutes):

Motivate Monte Carlo: “What if the NBA played 10,000 alternate versions of the season?”

Show that some correlations barely change (PTS–FGA), meaning they're structural

Show that some vary a lot (STL–BLK), meaning they're noisy

Explain the multivariate normal model used

Mention how this method is used in real sports analytics

This section alone can easily take 4–5 minutes if needed.
