---
title: "Draft Presentation"
author: "Ethan Schilling, Ivan Shokhrin, & Tyler Manuello"
date: "2025-12-01"
output:
  slidy_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Goal: Recreate confusion matrix of 2018 NBA data and compare to the 2022 season

- Use a confusion matrix to determine outcomes of predicting top player status
- Group top 15 players and create a heat map/confusion matrix to predict top player status
- Recreate heat mapped confusion matrix or a single row.
- Goal is to predict as many NBA All-Stars as possible.


## Their Confusion Matrix

```{r}
corr18 <- readRDS("corr18.rds")
corr18
```

## Our Confusion Matrix

```{r}
corr1 <- readRDS("corr1.rds")
corr1
```


## 
- Goal use a confusion matrix to determine outcomes of predicting top player status
- Group top 15 players and create a heat map/confusion matrix to predict top player status
- Recreate heat mapped confusion matrix or a single row.
- Goal is to predict as many NBA All-Stars as possible.

