title: "Final Project Presentation"
author: "Ben Eidler, Joseph Kroymann, Oriana Meldrum"
date: "12/7/2020"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Introduction: Home Field Advantage

- In sports psychology, the home field advantage is an acknowledged benefit. This benefit has been attributed to psychological effects supporting fans have on the competitors or referees; to psychological or physiological advantages of playing near home in familiar situations; to the disadvantages away teams suffer from changing time zones or climates, or from the rigors of travel; and in some sports, to specific rules that favor the home team directly or indirectly (“Home Advantage” 2020). However, due to the global COVID-19 pandemic, many sports leagues are playing without spectators. Does the absence of fans affect the predictions of the methods laid out in Lahvička’s 2015 paper?

## Introduction: Premier League

Premier League
- Soccer and stuff

## Literature 

Using Monte Carlo Simulation to Calculate Match Importance:
The Case of English Premier League
Jiřı́ Lahvička
- This article presents a new method of calculating match importance. Match
importance is defined as strength of relationship between the match result and a
given season outcome. Probabilities of all necessary match result-season outcome
combinations are estimated by Monte Carlo simulation. Using actual results of 12
seasons of English Premier League and betting odds, it is shown that both match
result and season outcome predictions are realistic. The method provides results
that are close to Jennett’s approach; however, it does not require ex post information and can be used for any type of season outcome.

## Data

In order to show the specified method is functional and accurate, we’ve pulled data from the English premier league official results website over the course of the 2019-2020 season, and the 2020-2021 season. The data consists of 20 different tables each corresponding to a specific team in the English Premier League. Of these teams, the three organizations which entered the league following relegation from the 2019-2020 season (Leeds United, Fulham Football Club, and West Bromwich Albion) have only 8 match dates worth of data. The remaining 17 teams have 19 match days worth of data spanning the previous two seasons, excluding Sheffield United who has played 20 games over that time period, and Manchester United who has only played 18 games over that time period. 

The collected data is fairly simple, and consists of 6 variables, excluding a column dedicated to indicating which team the table represents. The first three columns express the date of the home game for the specific team, the number of goals said team scored during that particular game, and the team's opponent on that given day. The next three columns express the date of the away game for the specific team, the number of goals said team surrendered during that particular game, and the team’s opponent on that given day. Whether or not a team actually wins a given game if not a factor in this specific analysis, which is why the data only includes single values for goals rather than multiple values to indicate victory or defeat.  An important factor in this data is the final number in the columns representing goals scored/goals surrendered, which indicates that teams average amount of goals scored at home or goals surrendered away. These values have been stored in an additional table dedicated to tracking the averages over this time span. 

The games we pulled data from were deliberately chosen because they were played during the COVID-19 global pandemic, and thus have been played in empty stadiums lacking any and all fanfare.   
 

```{r data, echo = TRUE}
#results_1920 <- results_1920 %>% select("Round Number", "Home Team", "Away Team", "Home", "Away")
#summary(results_1920)
```

## Method Step 1

Estimating Proabilities of Match results
- Expected goals for week 30 match: Chelsea vs Sheffield Utd: 

```{r Step 1, echo=FALSE }
library(ggplot2)
library(readr)
library(dplyr)
library(tidyverse)
#Step One
# importing game data for 19-20 season
results_1920 <- read.csv("data/19_20_PL_results.csv", nrows = 380)
results_1920 <- results_1920 %>% select(Round.Number, Home.Team, Away.Team, Home, Away)
# function that returns expected goals scored for a specific match (input two teams and a matchday)
expected_goals <- function(home, away, matchday) {
  rounds = seq((matchday - 1), (matchday - 19)) # matchdays to take data from
  
  home_data <- filter(results_1920, results_1920$Home.Team == home | results_1920$Away.Team == home)
  away_data <- filter(results_1920, results_1920$Home.Team == away | results_1920$Away.Team == away)
  
  home_scored <- 0
  home_conceded <- 0
  away_scored <- 0
  away_conceded <- 0
  
  for (i in rounds) {
    if (home_data[i,]$Home.Team == home) {
      home_scored = home_scored + home_data[i,]$Home
      home_conceded = home_conceded + home_data[i,]$Away
    }
    
    if (home_data[i,]$Away.Team == home) {
      home_scored = home_scored + home_data[i,]$Away
      home_conceded = home_conceded + home_data[i,]$Home
    }
    
    if (away_data[i,]$Away.Team == away) {
      away_scored = away_scored + away_data[i,]$Away
      away_conceded = away_conceded + away_data[i,]$Home
    }
    
    if (away_data[i,]$Home.Team == away) {
      away_scored = away_scored + away_data[i,]$Home
      away_conceded = away_conceded + away_data[i,]$Away
    }
  }
  
  home_expected <- (home_scored + away_conceded) / 2
  away_expected <- (away_scored + home_conceded) / 2
  
  c(home_expected/19, away_expected/19)
}
#kable(expected_goals("Chelsea", "Sheffield Utd", 30))
expected_goals("Chelsea", "Sheffield Utd", 30)
```
- WHY IS THIS SHOWING THE PACKAGE ERRORS TOO?

## Method Step 2

Monte Carlo Method for Prediction results

```{r Step 2}
p_home = function(goals_home, ev_home) {
  dpois(goals_home, ev_home)
}
p_away = function(goals_away, ev_away) {
  dpois(goals_away, ev_away)
}
mc_function = function (goals_home, ev_home, goals_away, ev_away){
  home_win_prob <- 0 
  away_win_prob <- 0
  tie_prob <- 0
  goals = seq(0, 9)
  
  # double for loop calculating probabilities for all scenarios up to 9 goals each (i = mc, j = mu)
  for (i in goals) {
    for (j in goals) {
      home <- p_home(goals_home, ev_home)
      away <- p_away(goals_away, ev_away)
      if (i == j) {
        tie_prob <- tie_prob + (home*away)
      }
      if (i > j) {
        home_win_prob <- home_win_prob + (home*away)
      }
      if (i < j) {
        away_win_prob <- away_win_prob + (home*away)
      }
    }
  }
  
  # numbers are all similar to the ones on pg. 395 of the paper
  #data.frame("mc win" = round(mc_win_prob, 3), "draw" = round(tie_prob, 3), "mu win" = round(mu_win_prob, 3))
  data.frame("Home win" = round(home_win_prob, 3), "draw" = round(tie_prob, 3), "Away win" = round(away_win_prob, 3)) 
}
#kable(mc_function( 1, 1.842, 0, 1.237))
mc_function( 1, 1.842, 0, 1.237)
```


## Method Step 3

Calculating Results-Outcome Association:
or who's going to get relegated if they fail

```{r Step 3}
```


## Verification

What are this week's results vs. What we predicted? 
- Should we make a table for this?

## Conclusion

Conclusion 
- Does the absence of fans affect the predictions of the methods laid out in Lahvička’s 2015 paper?

## Citations

Citations:

- “Home Advantage.” Wikipedia, Wikimedia Foundation, 14 Nov. 2020, en.wikipedia.org/wiki/Home_advantage.
- Lahvička, Jiří. “Using Monte Carlo Simulation to Calculate Match Importance: The Case of English Premier League.” Journal of Sports Economics, vol. 16, no. 4, 2015, pp. 390–409., doi:10.1177/1527002513490172.

